---
title: "Peer-graded Assignment: Course Project 1"
author: "JPLACLAU"
date: "January 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## 0.Creacion of wd " /ReproductableResearch "

```{r}

if(!file.exists("ReproductableResearch")){dir.create("ReproductableResearch")}



```


## 1.Code for reading in the dataset and/or processing the data

```{r}

# Download .zip file
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
f <- file.path(getwd(),"ReproductableResearch", "repdataFdataFactivity.zip")
download.file(url, f, mode="wb")
dateDownloaded <- date()
message("Archivo descargado en la fecha: ", dateDownloaded)

#unzip
unzip(path.expand("./ReproductableResearch/repdataFdataFactivity.zip"), files = NULL, list = FALSE, overwrite = TRUE,junkpaths = FALSE, exdir = ".", unzip = "internal",    setTimes = FALSE)

#read file
activity <-read.csv("./activity.csv")
head(activity)

```


## 2.Histogram of the total number of steps taken each day

```{r}

library(ggplot2)

#sum of steps per day
meantotalstepsperday <- tapply(activity$steps, activity$date, sum)

#histogram
qplot(meantotalstepsperday, geom="histogram",xlab = "Total number of steps taken each day", ylab ="Count of days") 

```


## 3.Mean and median number of steps taken each day

```{r}

#mean and median calculated with removed NA values

mean(meantotalstepsperday, na.rm = TRUE)
median(meantotalstepsperday, na.rm = TRUE)

```


## 4.Time series plot of the average number of steps taken

```{r}
# Calculating the mean of steps per interval
meanstepsperinterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)

plot(meanstepsperinterval, type = "l", main = ("Mean steps per 5 minutes interval on the day"), 
     ylab = "Number of steps", xlab = "5 min Interval of the day")


```


## 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}

activity[which.max(activity$steps), ]


```


## 6.Code to describe and show a strategy for imputing missing data

```{r}
#1.  total number of rows with NAs
missing <- is.na(activity$steps)
table(missing)

#2. Strategy to replace missing values: merging NA values with daily mean

#finding na positions
na_pos <- which(is.na(activity$steps))

#creating a vector of the means per interval
meansperinterval_vector <- rep(mean(activity$steps, na.rm = TRUE),times=length(na_pos))

#duplicating activity so original table stays intactc
filled_activity <- activity

#3. New dataset: replacing values
filled_activity[na_pos,"steps"] <- meansperinterval_vector
head(filled_activity)


```


## 7.Histogram of the total number of steps taken each day after missing values are imputed

```{r}
# sum of steps per day
sum_stepsperday <- aggregate(filled_activity$steps, by=list(filled_activity$date), FUN=sum)
head(sum_stepsperday)
names(sum_stepsperday) <- c("Date", "Total_steps")

#making histogram
hist(sum_stepsperday$Total_steps, 
     breaks=seq(from=0, to=25000, by=2500),
     col="purple", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Total number of steps taken each day (NA replaced by mean value of the day)")

#Mean and Median of filled values
mean(sum_stepsperday$Total_steps)
median(sum_stepsperday$Total_steps)


```

The impact of imputting missing data with daily means is that it makes the median closer to the mean and that is not necesarily accurate. Regarding the mean, there is no difference.


## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
#separating weekdays from weekends
weekday_or_weekend <- function(date) {
        day <- weekdays(date)
        if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
                return("weekday")
        else if (day %in% c("Saturday", "Sunday"))
                return("weekend")
        else
                stop("invalid date")
}

filled_activity$date <- as.Date(filled_activity$date)

#duplicating to avoid messing up already used data
New_filled_activity <- filled_activity

New_filled_activity$day <- sapply(filled_activity$date, FUN=weekday_or_weekend)

# Preparing means data weekdays/weekends
WdWe_means <- aggregate(steps ~ interval + day, data=New_filled_activity, mean)

# Compute the time serie plot
ggplot(WdWe_means, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
        xlab("5 min interval") + ylab("Average number of steps")


```


## 9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

```{r}

message("All code was 'echo=TRUE'. Thanks for reading")


```


